---
title: "ADA HW4"
author: "Liangquan Zhou lz2377"
date: "Thursday, October 02, 2014"
output: pdf_document
tags: hw4
---

## Problem 2 
From example 2 we know that $X_1, X_2, ... X_n \thicksim N(\mu_1, \sigma^2)$, $Y_1, Y_2, ... Y_n \thicksim N(\mu_2, \sigma^2)$, $\sigma = 10$, $\Delta= 4$, $\alpha = 0.05$ and $power = 0.8$

And the test is : $H_0: |\mu_1 - \mu_2 |=\Delta =0$ vs. $H_1: \Delta > 0$

a). Want to plot power as a function of sample size. Use normal approach to calculation the relation between power and sample size.

```{r, warning=FALSE,echo=TRUE}
power_n = function(sample_size, alpha, delta, sigma){
  
  z_alpha_2 = qnorm(1 - alpha/2)
  
  z_beta = (sqrt(sample_size) * delta) / (sqrt(2) * sigma) - z_alpha_2
  
  power = pnorm(z_beta)
  
  return(power)
}

# alpha = 0.05, sigma = 10, delta = 4
sample_size = seq(5, 100, by = 5)

power = power_n(sample_size, 0.05, 4, 10)

plot(sample_size, power, type="b", xlab="sample size", ylab="power", 
  main = "Sample Size v.s. Power")
```

b). For effect size $\Delta = \mu_1-\mu_2$, also use normal approach, and we can get the plot as follows:
```{r, warning=FALSE}
power_e = function(sample_size, alpha, beta, sigma){
  ## calculate delta
  z_alpha_2 = qnorm(1 - alpha/2)
  
  z_beta = qnorm(1 - beta)
  
  delta = sqrt(2) * (z_alpha_2 + z_beta) * sigma / sqrt(sample_size)
  
  return(delta)
}

# power = 0.8, beta = 1 - power = 0.2, alpha = 0.05, sigma = 10
sample_size = seq(5, 100, by = 5)

delta = power_e(sample_size, 0.05, 0.2, 10)

plot(sample_size, delta, type = "b", xlab = "sample size", ylab = "effective size", 
  main = "Sample Size v.s. Effective Size")
```

## Problem 3 
In this problem, $X \thicksim bin(n, p_1)$, $Y \thicksim bin(n, p_2)$, and $X$ and $Y$ are independent.  

The test is $H_0: p_1=p_2$ vs. $H_1: p_1 \neq p_2$. We want to do the same thing for proportion test. 

a). Want to plot the power as a function of sample size. 

1. Use normal approximation.

From the slides we know that 
$$
n = \frac{[Z_{\alpha/2} \sqrt{2 \bar{p} \bar{q}} + Z_\beta \sqrt{p_1 q_1 + p_2 q_2}]^2}{(p_2 - p_1)^2}
$$
and we can solve that
$$
Z_\beta = \frac{\sqrt{n(p_2-p_1)^2}-Z_{\alpha/2} \sqrt{2 \bar{p} \bar{q}}}{\sqrt{p_1q_1+p_2q_2}}
$$

```{r, warning=FALSE,echo=TRUE}
power_normal1 = function(sample_size, alpha, p1, p2){
  q1 = 1- p1
  
  q2 = 1- p2
  
  p_mean = (p1 + p2) / 2
  
  q_mean = 1 - p_mean
  
  z_alpha_2 = qnorm(1 - alpha/2)
  
  z_beta = (sqrt(sample_size * (p2 - p1)^2) - z_alpha_2 * sqrt(2 * p_mean * 
      q_mean)) / sqrt(p1 * q1 + p2 * q2)
  
  power = pnorm(z_beta)
  
  return(power)
}

# alpha = 0.05, p1 = 0.8, p2 = 0.9
sample_size = seq(5, 100, by = 5)

power = power_normal1(sample_size, 0.05, 0.8, 0.9)

plot(sample_size, power, type="b", xlab="sample size", ylab="power", main = 
    "Sample Size v.s. Power use Normal Approximation")
```

2. Use `arcsin transformation`

From the slides we know that
$$
n = \frac{(Z_{\alpha/2}+Z_\beta)^2}{\Delta^2}
$$
then 
$$
Z_\beta = \sqrt{n} \Delta - Z_{\alpha/2}
$$
```{r,warning=FALSE}
f = function(x) 2 * asin(sqrt(x))

power_arcsin1 = function(sample_size, alpha, p1, p2){
  
  delta = abs(f(p1) - f(p2))
  
  z_alpha_2 = qnorm(1 - alpha/2)
  
  z_beta = sqrt(sample_size) * delta - z_alpha_2
  
  power = pnorm(z_beta)
  
  return(power)
}

# alpha = 0.05, p1 = 0.8, p2 = 0.9
sample_size = seq(5, 100, by = 5)

power = power_arcsin1(sample_size, 0.05, 0.8, 0.9)
plot(sample_size, power, type = "b", xlab = "sample size", ylab = "power", 
  main = "Sample Size v.s. Power use arcsin Transformation")
```


b). Want to plot sample size as a function of effective size.

1. Use normal approximation.
```{r, warning=FALSE}
# fix p1 and let delta change, calculate sample size
power_normal2 = function(p1, p2, alpha, beta){
  
  q1 = 1- p1
  
  q2 = 1- p2
  
  delta = abs(p1 - p2)
  
  p_mean = (p1 + p2) / 2
  
  q_mean = 1 - p_mean
  
  z_alpha_2 = qnorm(1 - alpha/2)
  
  z_beta = qnorm(1- beta)
  
  sample_size = (z_alpha_2 * sqrt(2*p_mean*q_mean)+z_beta*sqrt(p1*q1+p2*q2))^2/(p2-p1)^2
  
  return(sample_size)
}

# set p2 > p1
p2 = seq(0.85, 1, by = 0.01)
p1 = 0.8
delta = p2 - p1
sample_size = power_normal2(p1, p2, 0.05, 1 - 0.8)

plot(delta, sample_size, type="b", main = "Delta v.s. Sample Size use Normal Approximation")
```

2. Use `arcsin transformation` 
```{r, message=FALSE}
f = function(x) 2 * asin(sqrt(x))

delta_arcsin = function(sample_size, alpha, beta){
  
  z_alpha_2 = qnorm(1 - alpha/2)
  
  z_beta = qnorm(1- beta)
  
  delta = (z_alpha_2 + z_beta)/sqrt(sample_size)
  
  return(delta)
}

# alpha = 0.05, beta = 1- power = 0.2
sample_size = seq(5, 100, by = 5)

delta = delta_arcsin(sample_size, 0.05, 0.2)
plot(delta,sample_size, type = "b", main = "Delta v.s. Sample Size use arcsin Transformation")
```
