---
title: "ADA HW2"
author: "Liangquan Zhou lz2377"
published: no
output: pdf_document
tags: hw2
---

Consider the data on Chicken Weights by Feed Type (`chickwts`) in R library `MASS`

## 1. Determine whether there is a significant difference in the mean weights of chicks fed meatmeal vs. those fed casein using each of the following procedures:

a). A parametric procedure

b). A non-parametric procedure

c). A re-sampling procedure

Discuss the assumption underlying each of the analyses, their validity, and any remedial measures to be taken.

### Solution
a). The hypothesis test:
$$
H_0: \mu_m- \mu_c=0 \ \ \ 
H_1: \mu_m-\mu_c \neq 0
$$
Assumptioms:

1. The weights of chicks fed on meatmeal and casein are independent.

2. The meatmeal and casein data are normally distributed.

3. The standard deviation of them are the same.

Check assumptions:
Use `Shapiro-Wilk Normality Test` to test the normality of two samples.

```{r ,message=FALSE,warning=FALSE}
library(MASS)
data(chickwts)
meatmeal=chickwts[chickwts[,2]=="meatmeal",1]
casein=chickwts[chickwts[,2]=="casein",1]
```

```{r,messaege=FALSE}
shapiro.test(meatmeal)
shapiro.test(casein)
```

And use `F-test` for standard deviations:
```{r}
var.test(meatmeal,casein)
```

Since these $p$-values are all greater than 0.05, we can think the assumption is reasionable.

And based on three assumptions above, we can use `two sample t-test`
```{r}
t.test(meatmeal,casein,var.equal=T)
```
From the result we can see that the $p$-value is greater than 0.05, so we do not reject $H_0$.


b). For a non-parametirc procedure, we use `Wilcoxon Rank-Sum Test`.
```{r}
wilcox.test(meatmeal,casein)
```
The $p$-value is also greater than 0.05, so we do not reject $H_0$.

c). Use bootstrap methods. Resampling `Meatmeal` and `Carsein` seperately, and for each iteration we compare the statistic with the statistic from the observed dataset.

```{r}
set.seed(1)
casein.bar= casein + mean(meatmeal) - mean(casein)
z.obs=(mean(meatmeal)-mean(casein))/sqrt(var(meatmeal)/length(meatmeal)+var(casein)/length(casein))
z.temp=c()
for(i in 1:1000){
  meatmeal.temp=sample(meatmeal,length(meatmeal),replace=T)
  casein.temp=sample(casein.bar,length(casein),replace=T)
  z.temp[i]=(mean(meatmeal.temp)-mean(casein.temp))/sqrt(var(meatmeal.temp)/length(meatmeal)+var(casein.temp)/length(casein))
}
p.value=sum(abs(z.temp)>=abs(z.obs))/1000
p.value
```
And the result is accord with the conclusion in (a) and (b). The $p$-value here is also greater than 0.05, then we should not reject $H_0$.


## 2. Using the data for chicks fed soybean and those on sunflower, compute the following, based on a suitable bootstrap method:

a). A 95% confidence interval for the difference in median weight for the two groups

b). A 95% bootstrap confidence interval for the ratio of the variances of soybean fed to sunflower fed chicks

c). A 95% confidence interval for the ratio of the variances of soybean fed to sunflower fed chicks

Discuss the assumption underlying each of the analyses, their validity, and any remedial measures to be taken

###Solution
a). The approach is to create a large number of samples from this pseudo-population using bootstrap. Then we use quantile function to find the 95% difference interval.
```{r,warning=FALSE}
soybean=chickwts[chickwts[,2]=="soybean",1]
sunflower=chickwts[chickwts[,2]=="sunflower",1]
library(bootstrap)
median.soy=bootstrap(soybean,1000,median)
median.sun=bootstrap(sunflower,1000,median)
quantile(abs(median.soy$thetastar-median.sun$thetastar),c(0.025,0.975))
```

b). Same method as (a).
```{r}
var.soy= bootstrap(soybean,1000,var)
var.sun = bootstrap(sunflower,1000,var)
quantile(var.soy$thetastar/var.sun$thetastar,c(0.025,0.975))
```

c). First test the normality and then use `var.test` to get the 95% confidence interval. 
```{r}
shapiro.test(soybean)
shapiro.test(sunflower)
var.test(soybean,sunflower)$conf.int
```
This means that the 95% confidence interval for the ratio of the variances of soybean fed to sunflower fed chicks is $[0.3622039, 3.9281145]$.

## 3. Assume that if the weight of a chick is below 258, that chick is classified under “LOW WEIGHT”. For chicks fed meatmeal vs. those fed soybean,

a)Determine whether there is a significant difference in the proportions of the chicks classified under “LOW WEIGHT”.

b)Construct a 95% confidence interval for the difference in the proportions of the chicks classified under “LOW WEIGHT”.

Discuss the assumption underlying each of the analyses, their validity, and any remedial measures to be taken.

###Solution
a) & b). The assumption is that the expected counts in `LOW WEIGHT` in both `meatmeal` and `soybean` are not too small (generally greater than 5). 

```{r}
n=length(meatmeal)+length(soybean) # total number
low.m=sum(meatmeal<258) 
low.s=sum(soybean<258)
p=(low.m+low.s)/n
min(p*length(meatmeal),(1-p)*length(meatmeal),p*length(soybean),(1-p)*length(soybean))
```
We can see that the expected counts of low weight or not low weight in`meatmeal` and `soybean` are all greater than 5. Then we can use `prop.test` to determine whether there is a significat difference in the proportions of the chicks classified under `LOW WEIGHT` and get its 95% confidence interval.
```{r}
prop.test(c(low.m,low.s),c(length(meatmeal),length(soybean)))
prop.test(c(low.m,low.s),c(length(meatmeal),length(soybean)))$conf.int
```
And from the result we can see that the $p$-value is greater than 0.05, thus we should not rejuct $H_0$. The 95% confidence interval is $[-0.7396 , 0.1812]$.